---
title: "BC Public Service Work Environment Survey 2015 Year-to-Year State of Engagement"
output: html_document
params:
  Organization: 
    label: "ORGID15"
    value: all
    input: select
    choices: [all, AG, AGRI, BCPSA, EBC, ENV, FIN, FLNR, GCPE, NGD, PGT, PO, PSSG, SDSI, TICS]
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(readr)
library(car)
library(pander)

knitr::opts_chunk$set(echo = TRUE)

#--------------- Set up the data for the RMD file ---------------

WES_data <- read_csv("finalMockData.csv", 
                          col_types = cols(COMMITMENT13 = col_number(), 
                                           COMMITMENT15 = col_number(),
                                           SAT13 = col_number(), 
                                           SAT15 = col_number()))

# Very basic input error checking
if (params$Organization != "all" &&
    !params$Organization %in% unique(WES_data$ORGID15)) {
  stop(paste0("The organization you provided is not in the data: ",
              params$Organization))
}

# Start the process to turn the data into what is needed to create a combined dataset

## Turn "Wide" data (where time is delineated by different variables) into "Long" data (where there is a variable
### for time called 'Year') 

longData <- WES_data %>% select(ORGID13,ORGANIZATION13,ORGID15,ORGANIZATION15, ENGSTATE13, ENGSTATE15) %>% 
  gather(Year, EngagementScore, ENGSTATE13:ENGSTATE15)

longData <- longData %>% mutate(Yr = ifelse(Year == "ENGSTATE13", "2013", "2015" ))

## Turn engagement score into a factor variable and recode to text labels

longData <- longData %>% mutate(EngagementScore = ifelse(Yr=="2013" & !is.na(ORGID13) & is.na(EngagementScore), "Non Respondents",
                                                         ifelse(Yr=="2015" & !is.na(ORGID15) & is.na(EngagementScore), "Non Respondents", EngagementScore)))

longData$EngagementScore <- factor(longData$EngagementScore,
                                   levels=
                                     c("Engaged", "Happily Detached", "Unhappily Dedicated", "Disengaged", "Non Respondents"))


## Loop to generate engagement state tables for all ministries:

for (i in unique(longData$ORGID15))
{
  assign(paste(i), longData %>% 
           filter(ORGID15==i | ORGID13==i) %>% 
           group_by(EngagementScore) %>% 
           summarize(Count2013 = sum(Yr=="2013" & ORGID13==i & !is.na(EngagementScore)), 
                     Count2015 = sum(Yr=="2015" & ORGID15==i & !is.na(EngagementScore))) %>% 
           mutate(NetGainOrLoss = Count2015 - Count2013) %>% 
           mutate(PercentageChange = ((Count2015-Count2013)/(Count2013))* 100))
}

all <- longData %>%
  group_by(EngagementScore) %>% 
  summarize(Count2013 = sum(Yr=="2013" & !is.na(EngagementScore)), 
                     Count2015 = sum(Yr=="2015" & !is.na(EngagementScore))) %>% 
           mutate(NetGainOrLoss = Count2015 - Count2013) %>% 
           mutate(PercentageChange = ((Count2015-Count2013)/(Count2013))* 100)

## Combine all the organization state tables very inelegantly

### Uncomment this when plugging in real data, and comment the proceeding block of code instead

# combinedStates <- bind_rows(all=all,AG=AG,AGRI=AGRI,ARR=ARR,AVED=AVED,BCPSA=BCPSA,
#                             CFD=CFD,CSCD=CSCD,EAO=EAO,EBC=EBC,EDUC=EDUC,EM=EM,
#                             EMBC=EMBC,ENV=ENV,FIN=FIN,FLNR=FLNR,GCPE=GCPE,HLTH=HLTH,
#                             JTSTL=JTSTL,MIT=MIT,NGD=NGD,OMB=OMB,PGT=PGT,PO=PO,
#                             PSSG=PSSG,SBRT=SBRT,SDSI=SDSI,TICS=TICS,TRAN=TRAN, .id="source")

combinedStates <- bind_rows(all=all,AG=AG,AGRI=AGRI,BCPSA=BCPSA,
                            EBC=EBC,ENV=ENV,FIN=FIN,FLNR=FLNR,GCPE=GCPE,NGD=NGD,
                            PGT=PGT,PO=PO,PSSG=PSSG,SDSI=SDSI,TICS=TICS, .id="source")

### Clean up the table, and append organizational names with the tables

combinedStates <- combinedStates %>% filter(!is.na(EngagementScore))
combinedStates <- rename(combinedStates, ORGID15 = source)

OrgNameID <- longData %>% filter(!is.na(ORGID15)) %>% group_by(ORGID15,ORGANIZATION15) %>% summarise()
OrgNameID <- rbind(as.data.frame(OrgNameID), c("all","all"))

combinedStates <- inner_join(combinedStates, OrgNameID %>% select(ORGANIZATION15, ORGID15), by = "ORGID15")
combinedStates <- combinedStates %>% mutate(PercentageChange = round(PercentageChange,digits=2))



IndexOrgName <- combinedStates %>% filter(ORGID15==params$Organization) %>% select(ORGANIZATION15) %>% distinct()
 

```

## `r IndexOrgName`

The purpose of this summary is to show how the state of engagement has changed in the organization between the 2013 and 2015 survey cycles

### How has the state of engagement changed?

The table below shows the overall number of employees in each state from both survey cycles.

```{r, echo = FALSE}


kable(combinedStates %>% filter(ORGID15==params$Organization), caption = "Table 1: Organizational Engagement State", digits = 2, 
      col.names = c("ID","Engagement", "Count 2013", "Count 2015", "Gain/Loss", "Change (%)", "Organization"), 
      align = c("l","l","c","c","c","c","l"))

# Creating a function to use for the paragraph below 

NetChange <- function(variable) 
{
combinedStates %>% filter(ORGID15==params$Organization) %>% filter(EngagementScore==variable) %>% select(NetGainOrLoss)
}

```


Since 2013, the organization has had a net `r ifelse((NetChange("Engaged"))>0,"gain of", ifelse((NetChange("Engaged"))<0,"loss of","change of"))` **`r abs((NetChange("Engaged")))`** engaged employees and a net `r ifelse((NetChange("Disengaged"))>0,"gain of", ifelse((NetChange("Disengaged"))<0,"loss of","change of"))` **`r abs((NetChange("Disengaged")))`** disengaged employees.
